---
title: "Covid91 vaccine study Final2022"
author: "Alondra Garcia"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_O$: There is no difference in the vaccine being effective in any of the subgroups.   
$H_A$: There is a difference in the vaccine and it being effective in some or most subgroups.


## Males

### Graphical Descriptive

```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")
```

According to the graph, the COVID-19 placebo had more of an effect than the vaccine having the advantage of 60% while the vaccine had roughly around 38%. An the 'OK' graph we can see that both placebo and vaccine had the same effect at around 50%.

### Numberical Descriptive

```{r}
table1 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

For the first treatment people took more of the placebo than the vaccine for COVID-19 with a leading total of 63.33 while the vaccine has 36.67. For the people who were infected by 'OK' sickness took about the same estimate on both the vaccine and placebo as we can see they are very close with a 50.10 on the placebo treatment and a 49.9 on the vaccine. For the second chart we could see that more people who were infected with the 'OK' virus took a lot more treatment of both placebo and vaccine as the numbers are more higher than COVID-19.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestM=fisher.test(table1)
ftestM
effecM=ftestM$estimate/(ftestM$estimate+1)*100
```

The Chi-square test shows that Male have a p-vaule of 0.00000000000026 which is less than 0.05, therefore we reject the null hypothesis 
For the fisher's exact test tells us that our p-vaule is 0.00000000000019.
the odds ratio is about 1.720532 and the efficacy is `r effecM`


## Females

### Graphical Descriptive

```{r}
barchartGC(~infected + sex,data=FinalData)
barchartGC(~infected + sex,data=FinalData, type="percent")
```

According to the graph there is no difference, both subgroups have about the same outcome.

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + sex, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestsex=fisher.test(table1)
ftestsex
effecsex=ftestsex$estimate/(ftestsex$estimate+1)*100
```

## LGBTQ

### Graphical Descriptive

```{r}
barchartGC(~infected + LGBTQ,data=FinalData)
barchartGC(~infected + LGBTQ,data=FinalData, type="percent")
```

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + LGBTQ, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestLGBTQ=fisher.test(table1)
ftestLGBTQ
effecLGBTQ=ftestLGBTQ$estimate/(ftestLGBTQ$estimate+1)*100
```

## Druggies

### Graphical Descriptive

```{r}
barchartGC(~infected + DrugUser,data=FinalData)
barchartGC(~infected + DrugUser,data=FinalData, type="percent")
```

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + DrugUser, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestDrugUser=fisher.test(table1)
ftestDrugUser
effecDrugUser=ftestDrugUser$estimate/(ftestDrugUser$estimate+1)*100
```

# Conclusion

In conclusion we reject the null hypothesis because according to most of the subgroup graphs there is a difference in the vaccine and it being effective in some or most subgroups.
