---
title: "Covid91 vaccine study Final2022"
author: "Alondra Garcia"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_O$: There is no difference in the vaccine being effective in any of the subgroups.   
$H_A$: There is a difference in the vaccine and it being effective in some or most subgroups.


## Males

### Graphical Descriptive

```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")
```

According to the graph, the COVID-19 placebo had more of an effect than the vaccine having the advantage of 60% while the vaccine had roughly around 38%. An the 'OK' graph we can see that both placebo and vaccine had the same effect at around 50%.

### Numberical Descriptive

```{r}
table1 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

For the first treatment people took more of the placebo than the vaccine for COVID-19 with a leading total of 63.33 while the vaccine has 36.67. For the people who were infected by 'OK' sickness took about the same estimate on both the vaccine and placebo as we can see they are very close with a 50.10 on the placebo treatment and a 49.9 on the vaccine. For the second chart we could see that more people who were infected with the 'OK' virus took a lot more treatment of both placebo and vaccine as the numbers are more higher than COVID-19.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestM=fisher.test(table1)
ftestM
effecM=ftestM$estimate/(ftestM$estimate+1)*100
```

The Chi-square test shows that Male have a p-value of 0.00000000000026 which is less than 0.05, therefore we reject the null hypothesis 
For the fisher's exact test tells us that our p-value is 0.00000000000019.
the odds ratio is about 1.720532 and the efficacy is `r effecM`


## Females

### Graphical Descriptive

```{r}
barchartGC(~infected + sex,data=FinalData)
barchartGC(~infected + sex,data=FinalData, type="percent")
```

According to the graph there is no difference, both subgroups have about the same outcome.

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + sex, data=FinalData)
rowPerc(table1)
colPerc(table1)
```
In the first treatment the female has more subjects who have taken the vaccine than the placebo. In the second treatment we can see that there is more female subjects that have taken the placebo than the vaccine.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestsex=fisher.test(table1)
ftestsex
effecsex=ftestsex$estimate/(ftestsex$estimate+1)*100
```
According to the chi-square test the Female p-value is about 0.6455 which is greater than 0.05, therefore we fail to reject the null hypothesis. 
For the fisher's exact test the p-value is 0.6386. The odds ratio is 0.9902268 and the efficacy is `r effecsex`

## LGBTQ

### Graphical Descriptive

```{r}
barchartGC(~infected + LGBTQ,data=FinalData)
barchartGC(~infected + LGBTQ,data=FinalData, type="percent")
```

According to the graph the straight category had a higher percentage than the gay category. The straight category had a high 80% while the gay category had less than a 20%.

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + LGBTQ, data=FinalData)
rowPerc(table1)
colPerc(table1)
```
According to the chart we can see that the first treatment had a lot more subjects, the straight category took more treatment than the gay category. As there is a significant difference in the straight category with around a 91%, while in the gay category only 8% took the treatment. On the second treatment both subgroups took about the same, they have both around 50%. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestLGBTQ=fisher.test(table1)
ftestLGBTQ
effecLGBTQ=ftestLGBTQ$estimate/(ftestLGBTQ$estimate+1)*100
```
The chi-square test shows that the p-value for this subgroup is 0.8924 which is greater than a 0.05 value, therefore we fail to reject the null hypothesis. For the fisher exact test it tells us that our p-value is 0.88, the odds ratio is 1.00584 and the efficacy is `r effecLGBTQ`

## Druggies

### Graphical Descriptive

```{r}
barchartGC(~infected + DrugUser,data=FinalData)
barchartGC(~infected + DrugUser,data=FinalData, type="percent")
```

According to the graph majority of the Drug users chose not to take the treatment. About a high 90% of them said no. While the rest of the subgroup said yes.

### Numberical Descriptive

```{r}
table1 <- xtabs(~treatment + DrugUser, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

According to the chart people who take drugs are far less likely to take the treatment on both vaccine or placebo. As you can see the percentage rate is very low, it's around 9% however people who don't take drugs are likely to accept treatment on both vaccine and placebo.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestDrugUser=fisher.test(table1)
ftestDrugUser
effecDrugUser=ftestDrugUser$estimate/(ftestDrugUser$estimate+1)*100
```
According to the chi-square test it shows that the p-value is 0.1811 which is higher than 0.05, therefore we fail to reject the null hypothesis. As for the fisher's exact test we have a p-value of 0.18, with a odds ratio of 1.052577 and with the efficacy of `r effecDrugUser`.

# Conclusion

In conclusion we reject the null hypothesis because according to most of the subgroup graphs there is a difference in the vaccine and it being effective in some or most subgroups. 
